title: "Lesson8"
author: "Maartje Holtslag & Jorn Habes"
date: "January 14, 2016"
output: html_document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries and download data
Load the required libraries, in this case the raster library.

```{r library}
library(raster)
```
Set the working directory and create a new folder called data.
```{r directory, results="hide"}
setwd("/home/user/git/geoScripting/Lesson8")
dir.create("./data")
```

Download the required data from github.
```{r download, results='hide'}
download.file(url = "https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/raw/gh-pages/data/GewataB1.rda", destfile= "data/GewataB1.rda", method = "wget")
download.file(url = "https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/raw/gh-pages/data/GewataB2.rda", destfile= "data/GewataB2.rda", method = "wget")
download.file(url = "https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/raw/gh-pages/data/GewataB3.rda", destfile= "data/GewataB3.rda", method = "wget")
download.file(url = "https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/raw/gh-pages/data/GewataB4.rda", destfile= "data/GewataB4.rda", method = "wget")
download.file(url = "https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/raw/gh-pages/data/GewataB5.rda", destfile= "data/GewataB5.rda", method = "wget")
download.file(url = "https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/raw/gh-pages/data/GewataB7.rda", destfile= "data/GewataB7.rda", method = "wget")
download.file(url = "https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/raw/gh-pages/data/vcfGewata.rda", destfile= "data/vcfGewata.rda", method = "wget")
download.file(url = "https://github.com/GeoScripting-WUR/AdvancedRasterAnalysis/raw/gh-pages/data/trainingPoly.rda", destfile= "data/trainingPoly.rda", method = "wget")
```
Load the required data from the folders.
```{r load, results='hide'}
load("data/GewataB1.rda")
load("data/GewataB2.rda")
load("data/GewataB3.rda")
load("data/GewataB4.rda")
load("data/GewataB5.rda")
load("data/GewataB7.rda")
load("data/vcfGewata.rda")
load("data/trainingPoly.rda")
```
##Raster analysis
Remove the cloud and water errors in the VCF data.
```{r removeerror}
vcfGewata[vcfGewata > 100] <- NA
``` 
Brick the Gewata raster data.
```{r brickGewata}
Gewata <- brick(GewataB1, GewataB2, GewataB3, GewataB4, GewataB5, GewataB7, vcfGewata)
GewataBands <- brick(GewataB1, GewataB2, GewataB3, GewataB4, GewataB5, GewataB7)
```
Create a data frame from the bricked Gewata rasters.
```{r dataframe}
GewataValues <- as.data.frame(getValues(Gewata))
```
Plot the Gewata pairs to compare to the VCF raster.
```{r pairing}
pairs(Gewata)
```
The plot above shows the correlation between the different rasters. The higher the correlation coefficient the more similar the data is they contain. For the VCF it seems that band 3 and 7 show the highest correlation with the VCF, and band 4 is not showing any correlation at all. This means that when trying to predict the tree cover based on these bands all but band 4 are usefull.

Create the correlation model and test whether the above mentioned statement is true.
```{r model}
correlationModel <- lm(vcf2000Gewata ~ gewataB1+ gewataB2 + gewataB3 + gewataB4 + gewataB5 + gewataB7, data = GewataValues) 
summary(correlationModel)
```
In the summary above the Pr (Probability) value determines whether the band has a predictive value for the model. The lower the Pr the better the band is to predict the tree cover. This means that all bands except band 7 have predictive value for the tree cover.

## Predicted tree cover raster and comparison
Create the predicted tree cover raster based on the correlation model and convert it to a data frame.
```{r predict}
predictTreeC <- predict(GewataBands, model=correlationModel, na.rm=TRUE)
predictTreeC[predictTreeC < 0] <- NA
predictTreeCdf <- as.data.frame(getValues(predictTreeC))
```
```{r predictplot, echo=FALSE}
plot(predictTreeC, main="Predicted tree cover", zlim=c(0,100))
```
Compare the predicted tree cover raster to the VCF raster.
```{r compare}
pairs(brick(predictTreeC, vcfGewata))
```
The plot above shows a very high correlation which means the model fits the actual raster pretty well. 

## Root Mean Squared Error
Calculate the RMSE (root mean squared error) to compare the predicted tree cover raster to VCF raster.
```{r RMSE}
RMSE <- (sqrt(mean((GewataValues$vcf2000Gewata-predictTreeCdf)^2, na.rm = T)))
``` 
```{r printRMSE, echo=FALSE}
RMSE
``` 
##Training classes comparison
Calculate the NDVI required for the training classes comparison.
```{r NDVI}
ndvi <- overlay(GewataB3, GewataB4, fun=function(x,y){(y-x)/(x+y)})
```
Make trainingPoly classes integers.
```{r integers}
trainingPoly@data$Code <- as.numeric(trainingPoly@data$Class)
```
Create the classes raster.
```{r classes}
classes <- rasterize(trainingPoly, ndvi, field='Code')
```
Calculate the RMSE for each class.
```{r calculate}
difference <- (vcfGewata-predictTreeC)^2
TreeCovDif <- zonal(difference, classes)
RMSEclasses <- (sqrt(TreeCovDif[,2]))
``` 
```{r RMSEhide, echo=FALSE}
RMSEclasses
```
The RMSE for the classes shows that the model predicts the tree cover best for forest (RMSE=4.99) and the worst for wetland (RMSE=11.91).